---
title: Summary of Cucurbita Habitat Modeling 
author: Daniel Carver. Colin Khoury 
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: yeti
    toc: yes
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("tmap")
# install.packages("sf")
# install.packages("DT")
# install.packages("XML")
# install.packages("scales")
library(knitr)
library(markdown)
library(rmarkdown)
library("tmap")
library("raster")
library("DT")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# function for flitering list based on character values 
include <- function (theList, toMatch){
  matches <- unique (grep(paste(toMatch,collapse="|"), 
                          theList, value=TRUE))
  return(matches)
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# taxa is the 
###DC
base_dr <- "C:/Users/danie/Desktop/aichiTest/aichiTest/gap_analysis"
taxa <- "Cucurbita_pepo_subsp._ovifera_var._ozarkana"

baseDir <- paste0(base_dr, '/', taxa)
csv <- list.files(baseDir, pattern = ".csv", recursive = TRUE , full.names = TRUE)
```


### Summary Statistics for Model Runs for `r taxa`

This table show the statistical results of all median models. 
```{r echo=FALSE, message=FALSE, warning=FALSE}

dfSum <- data.frame()
runs <- list.dirs(baseDir,recursive = FALSE , full.names = FALSE)
metrics <- include(csv, "eval_metrics.csv")
for(i in 1:length(runs)){
  value <- grepl(runs[i], metrics[i], fixed = TRUE)
  if(any(value)){
  dt1 <- read.csv(metrics[i])
  dt1$run <- runs[i]
  dfSum <- rbind(dfSum, dt1)
  }else{
    next
  }
}

DT::datatable(dfSum)
```

### All Models Addative Map
This map combines all threshold maps into one. The darkest areas are where all models agree that the location represents potential suitable habitat.  
```{r echo=FALSE, message=FALSE, warning=FALSE}
tif <- list.files(path = baseDir, pattern = '.tif', recursive = TRUE, full.names = TRUE)
thres <- include(tif, "spdist_thrsld")
rasters <- lapply(thres,raster)
size <- lapply(rasters, ncell)
maxRas <- which.max(size)

ras1 <- rasters[maxRas]
ras1 <- ras1[[1]]
rastersThin <- rasters[-c(maxRas)]

ras1[is.na(ras1[])] <- 0 


#build the first raster here 
if(length(thres) > 0 ){
  for(i in 1:length(rastersThin))
    {
    rast2 <- grepl(runs[i], metrics, fixed = TRUE)
    if(any(rast2))
      {
      ras <- raster(thres[i])
      ras[is.na(ras[])] <- 0 
      ras <- extend(ras, extent(ras1), value = 0)
      ras1 <- ras1 + ras
      }else
    {
    next
    }
  }
}else{
  ras1 <- NA
}

ras1[ras1[] <= 0] <- NA 




tmap_mode("view")
if(class(ras1) == "RasterLayer"){
  qtm(shp = ras1)
}else{
  print("no rasters are present")
}
```
This map above is composed of `r length(rasters)` model results. 


### Biome and Ecoregion Maps

The map below shows the combination map of our two most conservative models. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
for(class in rasters){
  bio <- grepl("bio", metrics, fixed = TRUE)
  bioRast <- rasters[bio][[1]]
}

for(class in rasters){
  eco <- grepl("ecoR", metrics, fixed = TRUE)
  ecoRast <- rasters[eco][[1]]
}

# for(class in rasters){
#   ecoV <- grepl("ecoV", metrics, fixed = TRUE)
#   ecoRast1 <- rasters[ecoV][[1]]
# }


bio1 <- bioRast
eco1 <- ecoRast
# ecoV <- ecoRast1

ecoFull <- extend(eco1, extent(bio1), value = 0)
# ecoFullv <- extend(ecoV, extent(bio1), value = 0)

eco1[is.na(eco1[])] <- 0 
bio1[is.na(bio1[])] <- 0 
# ecoV[is.na(ecoV[])] <- 0 
# ecoBio <- ecoFull + bioRast + ecoFullv
ecoBio <- ecoFull + bio1
if(class(ecoBio) == "RasterLayer"){
  qtm(shp = ecoBio)
}else{
  print("no rasters are present")
}
```




### EcoRegion 
The map below is an example of the prediction map that is limited to ecoregions in which known presense locations are found. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
if(class(ecoRast) == "RasterLayer"){
  qtm(shp = ecoRast)
}else{
  print("no rasters are present")
}

```


### Biome 
The map below is an example of the prediction map that is limited to biomes in which known presense locations are found. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(class(bioRast) == "RasterLayer"){
  qtm(shp = bioRast)
}else{
  print("no rasters are present")
}
```



### Ecoregion Nipals with VifStep Correlation test 
Map for the principal component  and correlation removal variable selection proces. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
for(class in rasters){
  testr <- grepl("nipalsvirStep", metrics, fixed = TRUE)
    if(any(testr)){
    testr <- rasters[testr][[1]]
  }else{
  print("no rasters are present")
  }
}


if(class(testr) == "RasterLayer"){
  qtm(shp = testr)
}else{
  print("no rasters are present")
}



```


### Ecoregion rfUtilities with Correlation test 
Map for the rfUtilities with the removal of correlated variables.  
```{r echo=FALSE, message=FALSE, warning=FALSE}
for(class in rasters){
  testr <- grepl("rfUtilities", metrics, fixed = TRUE)
    if(any(testr)){
    testr <- rasters[testr][[1]]
  }else{
  print("no rasters are present")
  }
}


if(class(testr) == "RasterLayer"){
  qtm(shp = testr)
}else{
  print("no rasters are present")
}


```


### Ecoregion Vsurf threshold with Correlation test 
Map for the vsurf threshold selection with correlation test selection proces. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
for(class in rasters){
  testr <- grepl("vsurfThres", metrics, fixed = TRUE)
    if(any(testr)){
    testr <- rasters[testr][[1]]
  }else{
  print("no rasters are present")
  }
}


if(class(testr) == "RasterLayer"){
  qtm(shp = testr)
}else{
  print("no rasters are present")
}



```

